\documentclass[11pt,a4paper]{article}

% --- Encoding & fonts ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}            % Better quality scalable fonts

% --- Page geometry ---
\usepackage[
    a4paper,
    top=2.5cm,
    bottom=2.5cm,
    left=3cm,
    right=3cm,
    headheight=14pt
]{geometry}

% --- Spacing ---
\usepackage{setspace}
\onehalfspacing

% --- Math ---
\usepackage{amsmath,amssymb,amsfonts}

% --- Graphics ---
\usepackage{graphicx}
\usepackage{float}
\usepackage{tikz}

% --- Tables ---
\usepackage{booktabs}
\usepackage{array}

% --- Code listings ---
\usepackage{listings}
\usepackage{xcolor}

% --- Headers and footers ---
\usepackage{fancyhdr}

% --- Hyperlinks (load late, before cleveref) ---
\usepackage[
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=blue,
    bookmarks=true,
    pdftitle={LectureMind: AI-Powered Lecture Video Analysis and Summarization Platform},
    pdfauthor={Group Report},
    pdfsubject={COMP5575 Group Project Report}
]{hyperref}

% --- Typography ---
\usepackage{microtype}

% --- Paragraph formatting ---
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

% -------------------------------------------------------
% Code listing style
% -------------------------------------------------------
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.97,0.97,0.95}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{blue}\bfseries,
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2,
    xleftmargin=12pt,
    xrightmargin=0pt,
    frame=single,
    rulecolor=\color{gray!40},
    framesep=4pt
}
\lstset{style=mystyle}

\lstdefinelanguage{json}{
    basicstyle=\ttfamily\footnotesize,
    columns=fullflexible,
    keepspaces=true,
    breaklines=true,
    showstringspaces=false,
    string=[s]{"}{"},
    stringstyle=\color{codepurple},
    morecomment=[l]{//},
    morecomment=[s]{/*}{*/},
    morestring=[b]"
}

\lstdefinelanguage{text}{
    basicstyle=\ttfamily\footnotesize,
    columns=fullflexible,
    keepspaces=true,
    breaklines=true,
    showstringspaces=false
}

% -------------------------------------------------------
% Page style
% -------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textit{LectureMind: AI-Powered Lecture Video Analysis}}
\fancyhead[R]{\small COMP5575 Group Project}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Plain style for title page / first page
\fancypagestyle{plain}{
    \fancyhf{}
    \fancyfoot[C]{\thepage}
    \renewcommand{\headrulewidth}{0pt}
}

% -------------------------------------------------------
% Section title formatting
% -------------------------------------------------------
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}[\titlerule]
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalsize\itshape\bfseries}{\thesubsubsection}{1em}{}
\titlespacing*{\section}{0pt}{16pt}{8pt}
\titlespacing*{\subsection}{0pt}{12pt}{6pt}
\titlespacing*{\subsubsection}{0pt}{10pt}{4pt}

% -------------------------------------------------------
% Caption formatting
% -------------------------------------------------------
\usepackage[
    font=small,
    labelfont=bf,
    labelsep=period
]{caption}

% -------------------------------------------------------
% Document
% -------------------------------------------------------
\begin{document}

% ===== Title Page =====
\begin{titlepage}
    \centering
    \vspace*{2cm}

    {\Large \textbf{The Hong Kong Polytechnic University}}\\[0.4cm]
    {\large Department of Computing}\\[1.5cm]

    \rule{\linewidth}{0.5pt}\\[0.4cm]
    {\LARGE \bfseries LectureMind: AI-Powered Lecture Video Analysis\\[0.3cm] and Summarization Platform}\\[0.4cm]
    \rule{\linewidth}{0.5pt}\\[1.5cm]

    {\large \textbf{COMP5575 -- Group Project Report}}\\[2cm]

    \begin{tabular}{ll}
        \textbf{Group Members:} & Author A \\
                                & Author B \\
                                & Author C \\[1cm]
        \textbf{Course:}        & COMP5575 \\[0.4cm]
        \textbf{Semester:}      & 2025 Fall \\[0.4cm]
        \textbf{Submission:}    & \today \\
    \end{tabular}

    \vfill

    {\small \textit{Submitted in partial fulfilment of the requirements for COMP5575}}
\end{titlepage}

% ===== Abstract =====
\newpage
\thispagestyle{plain}
\begin{center}
    {\large \bfseries Abstract}
\end{center}
\noindent
Lecture videos are a fundamental component of modern education, yet students face significant challenges in efficiently navigating, reviewing, and extracting knowledge from lengthy lecture recordings. This report presents LectureMind, an advanced AI-powered platform that transforms raw lecture videos into structured, searchable, and interactive learning resources through a multi-stage processing pipeline.

Our system employs a directed acyclic graph (DAG) of ten asynchronous tasks that process videos through: (1) SSIM-based slide transition detection using multithreaded structural similarity analysis, (2) automatic speech recognition via Alibaba DashScope Qwen3-ASR, (3) hybrid video chunking combining slide changes and silence gaps, (4) dual-resolution thumbnail generation (200px for web display, 1920px for OCR), (5) slide OCR using a multimodal LLM on high-resolution thumbnails, (6) fine-grained knowledge point extraction, (7) vector embedding into ChromaDB, (8) coarse-grained lecture summarization, and (9--10) hierarchical mindmap generation.

The core innovation lies in our three-mode Retrieval-Augmented Generation (RAG) system: LLM Direct (no retrieval), Fast RAG (single-pass vector retrieval with quality-gated fallback), and Agentic RAG (multi-step LangGraph ReAct agent with six tools including slide-OCR search). A cascading fallback chain (Agentic RAG to Fast RAG to LLM Direct) ensures reliable answers under all conditions. An automated LLM-as-judge evaluation framework quantifies performance across question types: Agentic RAG achieves the highest average score (74.2/100) while Fast RAG maintains zero hallucination rate (0\%) through conservative retrieval thresholds. The entire system is containerised with Docker Compose and fully configurable via environment variables.

\bigskip
\noindent\textbf{Keywords:} Educational Technology, Video Analysis, Retrieval-Augmented Generation, Large Language Models, Speech Recognition, Slide OCR

% ===== Table of Contents =====
\newpage
\tableofcontents
\listoftables

% ===== Chapters =====
\newpage
\input{chapters/01_introduction}
\input{chapters/02_related_work}
\input{chapters/03_architecture}
\input{chapters/04_preprocessing}
\input{chapters/05_chunking}
\input{chapters/06_knowledge}
\input{chapters/07_rag}
\input{chapters/08_implementation}
\input{chapters/09_evaluation}
\input{chapters/10_conclusion}

% ===== References =====
\newpage
\begin{thebibliography}{99}

\bibitem{guo2014video}
Guo, P.\,J., Kim, J., \& Rubin, R. (2014).
How video production affects student engagement: An empirical study of MOOC videos.
\textit{Proceedings of the 1st ACM Conference on Learning@Scale}, 41--50.

\bibitem{copeland2009lecture}
Copeland, S., \& Jones, R. (2009).
Lecture capture: A new approach to teaching and learning?
\textit{Proceedings of the 7th International Conference on Information Technology Based Higher Education and Training}, 512--517.

\bibitem{cui2016automatic}
Cui, Y., Hu, Y., \& Chen, W. (2016).
Automatic segmentation of educational videos based on slide transition detection.
\textit{IEEE International Conference on Multimedia and Expo}, 1--6.

\bibitem{pan2020educational}
Pan, Y., et al. (2020).
Educational video segmentation using multimodal deep learning.
\textit{IEEE Transactions on Learning Technologies}, 13(4), 789--801.

\bibitem{chen2023llm}
Chen, X., et al. (2023).
LLM-based automatic quiz generation from lecture transcripts.
\textit{Proceedings of the ACM Conference on Learning@Scale}, 112--124.

\bibitem{lewis2020retrieval}
Lewis, P., et al. (2020).
Retrieval-augmented generation for knowledge-intensive NLP tasks.
\textit{Advances in Neural Information Processing Systems}, 33, 9459--9474.

\bibitem{lee2021textbook}
Lee, J., \& Kim, S. (2021).
Textbook question answering using retrieval-augmented generation.
\textit{Proceedings of the NAACL-HLT}, 2345--2356.

\bibitem{wang2022lecture}
Wang, H., et al. (2022).
LectureQA: A dataset and model for question answering on lecture videos.
\textit{Proceedings of the ACL}, 4567--4580.

\bibitem{xu2017video}
Xu, D., et al. (2017).
Video question answering via gradually refined attention over appearance and motion.
\textit{Proceedings of the ACM Multimedia}, 1645--1653.

\bibitem{li2023zero}
Li, J., et al. (2023).
Zero-shot video question answering using pretrained language models.
\textit{Proceedings of the CVPR}, 12345--12356.

\bibitem{hoffmann2011knowledge}
Hoffmann, R., et al. (2011).
Knowledge-based weak supervision for information extraction of overlapping relations.
\textit{Proceedings of the ACL-HLT}, 541--550.

\bibitem{gutierrez2023llm}
Gutierrez, B., et al. (2023).
LLM-based concept extraction from educational texts.
\textit{Proceedings of the EDM}, 234--245.

\bibitem{pan2023ontology}
Pan, S., et al. (2023).
Ontology construction using large language models.
\textit{Proceedings of the ISWC}, 567--584.

\bibitem{wang2004image}
Wang, Z., Bovik, A.\,C., Sheikh, H.\,R., \& Simoncelli, E.\,P. (2004).
Image quality assessment: from error visibility to structural similarity.
\textit{IEEE Transactions on Image Processing}, 13(4), 600--612.

\bibitem{reimers2019sentence}
Reimers, N., \& Gurevych, I. (2019).
Sentence-BERT: Sentence embeddings using Siamese BERT-networks.
\textit{Proceedings of the EMNLP-IJCNLP}, 3982--3992.

\end{thebibliography}

\end{document}
