\section{Hybrid Video Chunking Algorithm}
\label{sec:chunking}

\subsection{Motivation}

Accurate video segmentation is crucial for downstream knowledge extraction. Simple segmentation strategies have limitations:

\begin{itemize}
    \item \textbf{Slide-only segmentation:} May produce sections that are too short or split coherent topics across multiple slides
    \item \textbf{Time-based segmentation:} Fixed-duration chunks ignore content boundaries, splitting explanations mid-sentence
    \item \textbf{Silence-only segmentation:} Natural pauses don't always correspond to topic boundaries
\end{itemize}

Our hybrid approach combines multiple signals to produce topically coherent sections.

\subsection{Signal Sources}

\subsubsection{Slide Change Candidates}

Slide transitions from SSIM detection provide strong physical boundaries. Let $S = \{s_1, s_2, ..., s_n\}$ be the set of slide change timestamps.

\subsubsection{Silence Gap Candidates}

Silence gaps in speech indicate natural pause points where topic transitions often occur. From the ASR transcript, we identify gaps between consecutive sentences:

\begin{equation}
    G = \{g_i : \text{begin}_{i+1} - \text{end}_i \geq \tau_{silence}\}
\end{equation}

where $\tau_{silence} = 2.0$ seconds. For each qualifying gap, the midpoint is used as a candidate:

\begin{equation}
    t_{silence} = \frac{\text{end}_i + \text{begin}_{i+1}}{2}
\end{equation}

\subsubsection{Semantic Similarity (Optional)}

Semantic similarity analysis detects topic shifts in continuous speech where no silence gap exists. Using sentence-transformers \cite{reimers2019sentence}, we compute cosine similarity between text windows before and after each candidate point:

\begin{equation}
    \text{sim}(c) = \cos(E(W_{before}), E(W_{after}))
\end{equation}

where $E$ is the embedding function (all-MiniLM-L6-v2) and $W$ represents a 30-second text window. Candidates with similarity above threshold ($\tau_{semantic} = 0.5$) are rejected as they indicate topic continuity.

This feature is currently disabled in production to reduce memory usage on constrained systems (the embedding model requires ~200MB RAM when loaded).

\subsection{Chunking Algorithm}

Listing~\ref{alg:hybrid-chunking} presents the complete hybrid chunking process.

\begin{lstlisting}[language=Python, caption={Hybrid Video Chunking}, label={alg:hybrid-chunking}]
Input: Slide changes S, silence gaps G, min duration d_min
Output: List of section boundaries B

C = S union G          // Merge candidates
C = sort(C)
B = {0.0}              // Start with video beginning
last = 0.0

for each candidate c in C:
    if c - last >= d_min:
        if semantic check disabled OR sim(c) < tau_semantic:
            B = B union {c}
            last = c

B = B union {t_video_end}  // Add video end
return sections from consecutive pairs in B
\end{lstlisting}

\subsection{Section Creation}

For each chunk $(t_{start}, t_{end})$, a \texttt{VideoSection} record is created:

\begin{itemize}
    \item Extract transcript text by concatenating all sentences within the time range
    \item Find the closest thumbnail to $t_{start}$ for visual representation
    \item Assign initial title "Section N" (later updated by LLM in Task 6)
    \item Store begin/end times and transcript text
\end{itemize}

\begin{table}[H]
    \centering
    \caption{Hybrid Chunking Parameters}
    \begin{tabular}{@{}lll@{}}
        \toprule
        \textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \midrule
        $d_{min}$ & 30.0 s & Minimum section duration \\
        $\tau_{silence}$ & 2.0 s & Silence gap threshold \\
        $\tau_{semantic}$ & 0.5 & Semantic similarity threshold \\
        semantic check & disabled & Optional feature flag \\ \bottomrule
    \end{tabular}
\end{table}

\subsection{Example Output}

For a typical 60-minute lecture:

\begin{itemize}
    \item Slide changes detected: 12-18
    \item Silence gaps identified: 8-15
    \item Final sections created: 8-12
    \item Average section duration: 5-8 minutes
\end{itemize}

The hybrid approach typically reduces the raw candidate count by 30-40\% through minimum duration filtering, producing sections that align well with topical boundaries.
