\section{Introduction}

\subsection{Background}

Lecture videos have become a standard medium for higher education, online learning, and professional training. A single course may contain many hours of recorded material, often accompanied by slides, spoken explanations, and informal examples. While this abundance improves access, it also creates a navigation problem: students rarely need the entire video at once. More often, they want to locate a definition, revisit an explanation, compare related concepts, or ask a targeted question about a specific lecture segment. Conventional video players offer only linear playback and coarse time-based seeking, which makes content review inefficient for knowledge-intensive subjects.

This problem is especially visible in technical lectures. Important information is distributed across multiple modalities: the instructor's speech introduces concepts, the slides encode structure, and longer explanations may extend across several minutes without any obvious boundaries. A useful educational assistant must therefore do more than transcribe audio. It must segment the lecture into meaningful units, identify the knowledge contained in each unit, organize those units into a coherent structure, and support question answering grounded in lecture evidence.

\subsection{Problem Definition}

This report addresses the following problem: given a raw lecture video, how can we transform it into a structured and interactive learning resource that supports efficient review and accurate question answering? The goal is not only to summarize the video, but to bridge three levels of understanding:

\begin{itemize}
\item the \textit{temporal level}, where the system identifies meaningful lecture boundaries;
\item the \textit{semantic level}, where the content is converted into explicit knowledge units;
\item the \textit{interactive level}, where users can query the lecture and obtain grounded answers.
\end{itemize}

To solve this problem, the system must handle noisy speech, sparse slide transitions, and the mismatch between lecture delivery and knowledge structure. A good segmentation method should not rely on a single cue, because slide changes alone are too coarse and silence alone is too unstable. Likewise, a question answering system should not depend on a single reasoning strategy, because some questions require direct recall while others demand retrieval or multi-step exploration.

\subsection{Method Overview}

LectureMind is designed as a multi-stage algorithmic framework for lecture video understanding. The pipeline begins with speech transcription and visual analysis, then combines these signals to produce semantically meaningful lecture chunks. Each chunk is further processed to extract fine-grained knowledge points, which are later aggregated into higher-level summaries and a mindmap-like structure. On top of these structured representations, the system provides three question answering modes: direct answering without retrieval, single-pass retrieval-augmented generation, and agentic multi-step retrieval.

Figure~\ref{fig:system-overview} summarizes this end-to-end transformation from raw video input to interactive educational output.

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
        node distance=1.5cm,
        box/.style={rectangle, draw, rounded corners, minimum width=3cm, minimum height=1cm, align=center},
        arrow/.style={->, >=stealth, thick}
    ]
        \node[box, fill=blue!10] (input) {Video Input};
        \node[box, fill=green!10, below of=input] (preprocess) {Preprocessing\\ASR | HLS | SSIM};
        \node[box, fill=yellow!10, below of=preprocess] (chunking) {Hybrid Chunking\\Slide + Silence + Semantic};
        \node[box, fill=orange!10, below of=chunking] (knowledge) {Knowledge Extraction\\Fine-grained | Coarse-grained};
        \node[box, fill=purple!10, below of=knowledge] (storage) {Vector Database\\ChromaDB};
        \node[box, fill=red!10, below of=storage] (query) {RAG Query Engine\\Quick RAG | Agent};
        \node[box, fill=cyan!10, below of=query] (output) {Interactive Output\\Chat | Mindmap | Citations};

        \draw[arrow] (input) -- (preprocess);
        \draw[arrow] (preprocess) -- (chunking);
        \draw[arrow] (chunking) -- (knowledge);
        \draw[arrow] (knowledge) -- (storage);
        \draw[arrow] (storage) -- (query);
        \draw[arrow] (query) -- (output);
    \end{tikzpicture}
    \caption{LectureMind System Architecture Overview}
    \label{fig:system-overview}
\end{figure}

The central idea of the framework is that lecture understanding should be hierarchical. A lecture is first decomposed into segments, each segment is mapped to explicit knowledge, and the resulting knowledge is organized so that retrieval and reasoning can operate on units that are smaller than the whole lecture but richer than raw transcript sentences. This hierarchy makes it possible to support both efficient access and evidence-aware responses.

\subsection{Contributions}

The main contributions of this report are threefold.

\begin{itemize}
\item We present an end-to-end method for transforming lecture videos into structured learning resources, covering segmentation, knowledge extraction, and grounded question answering.
\item We propose a hybrid chunking strategy that combines slide transition cues with speech-based boundaries, improving robustness over single-signal segmentation.
\item We introduce a three-mode retrieval-augmented answering framework that balances reasoning depth, response speed, and answer reliability through a cascading fallback design.
\end{itemize}